---
title: "Harry Potter Sentiment Analysis"
author: "Georgianna James"
date: '2022-03-09'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up

## Required Packages 

```{r message=FALSE, warning=FALSE}
library(harrypotter)
library(tidyverse)
library(tidytext)
library(textdata)

theme_set(theme_minimal())
```



## Creating Data Frame 

Bradley Boehmke's ```harrypotter``` package contains the text of the entire seven book Harry Potter series. The package consists of seven data sets, one for each novel, with the text stored as one character vector with each element representing a single chapter. 

For my sentiment analysis, I combine all seven novels and store each word as a single token associated with the chapter and title of the book it is from. In order to do this, first, I created a character vector for the title of each book. Then I combined all of the individual novel data sets into a list and named them using my title character vector. Next, I make each book its own tibble and and merged them into ```book``` using the ```map_df()``` function. Then I converted ```book``` to a factor with a level for each Harry Potter novel. Because the original data contained a row for each chapter, I created a chapter column by grouping by book and setting ```chapter``` equal to ```row_number(book)```. Finally, I unnested the dataframe, making each word its own token. 




```{r}
# character vector of titles of all books

hp_books <- c("philosophers_stone", "chamber_of_secrets",
              "prisoner_of_azkaban", "goblet_of_fire",
              "order_of_the_phoenix", "half_blood_prince",
              "deathly_hallows")
```


```{r}
# combine book data sets into a list
hp_words <- list(
  philosophers_stone,
  chamber_of_secrets,
  prisoner_of_azkaban,
  goblet_of_fire,
  order_of_the_phoenix,
  half_blood_prince,
  deathly_hallows
) %>%
  # name each list element
  set_names(hp_books) %>%
  # convert each book to a data frame and merge into a single data frame
  map_df(as_tibble, .id = "book") %>%
  # convert book to a factor
  mutate(book = factor(book, levels = hp_books)) %>%
  # remove empty chapters
  drop_na(value) %>%
  # create a chapter id column
  group_by(book) %>%
  mutate(chapter = row_number(book)) %>%
  ungroup() %>%
  # tokenize the data frame
  unnest_tokens(word, value)

head(hp_words)
```



## Visualization of Most Frequent Words, by book




```{r most frequent words, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=12}
hp_words %>%
  # delete stopwords
  anti_join(stop_words) %>%
  # summarize count per word per book
  count(book, word) %>%
  # get top 15 words per book
  group_by(book) %>%
  slice_max(order_by = n, n = 15) %>%
  mutate(word = reorder_within(word, n, book)) %>%
  # create barplot
  ggplot(aes(x = word, y = n, fill = book)) + 
  geom_col() +
  scale_x_reordered() +
  labs(title = "Most frequent words in Harry Potter",
       x = NULL,
       y = "Word count") +
  facet_wrap(facets = vars(book), scales = "free") +
  coord_flip() +
  theme(legend.position = "none")
```



# Sentiment Analysis Using Bing Dictionary 

### Generate data frame with sentiment derived from the Bing dictionary

```{r}
# creating df from bBng dictionary
hp_sentiment_analysis <- get_sentiments("bing")

# joining by words from hp books that are also in the Bing dictionairy

harry_potter_sentiment_analysis <- left_join(hp_words, hp_sentiment_analysis, by = "word") %>% 
  drop_na(sentiment)
```


## Most frequently used sentimental words, both positive and negative, in the whole series

```{r pos and neg words in series, echo=FALSE}
# all series
harry_potter_sentiment_analysis %>% 
  # generate frequency count for each word and sentiment
  group_by(sentiment) %>%
  count(word) %>%
  # extract 10 most frequent pos/neg words
  group_by(sentiment) %>%
  slice_max(order_by = n, n = 10) %>%
  # prep data for sorting each word independently by facet
  mutate(word = reorder_within(word, n, sentiment)) %>%
  # generate the bar plot
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  # used with reorder_within() to label the axis tick marks
  scale_x_reordered() +
  facet_wrap(facets = vars(sentiment), scales = "free_y") +
  labs(
    title = "Sentimental words used in the Harry Potter series",
    x = NULL,
    y = "Number of occurences in all seven books"
  ) +
  coord_flip()
```



### Generate a data frame that counts the frequency of each word by book and sentiment. 



```{r}

hp_pos_neg_book <- harry_potter_sentiment_analysis %>%
  # count freq for each book, word, and sentiment
  group_by(book, sentiment) %>%
  count(word) %>%
  # extract 10 most frequent pos/neg words per book
  group_by(book, sentiment) %>%
  slice_max(order_by = n, n = 10)
```



## Most Frequently Used Positive Words, by book

```{r echo=FALSE}
## positive words
hp_pos_neg_book %>%
  filter(sentiment == "positive") %>%
  mutate(word = reorder_within(word, n, book)) %>%
  ggplot(aes(word, n, fill = book)) +
  geom_col(show.legend = FALSE) +
  scale_x_reordered() +
  facet_wrap(facets = vars(book), scales = "free_y") +
  labs(
    title = "Positive words used in the Harry Potter series",
    x = NULL,
    y = "Number of occurences"
  ) +
  coord_flip()
```



## Most Frequently Used Negative Words, by book


```{r echo=FALSE}
## negative words
hp_pos_neg_book %>%
  filter(sentiment == "negative") %>%
  mutate(word = reorder_within(word, n, book)) %>%
  ggplot(aes(word, n, fill = book)) +
  geom_col(show.legend = FALSE) +
  scale_x_reordered() +
  facet_wrap(facets = vars(book), scales = "free_y") +
  labs(
    title = "Negative words used in the Harry Potter series",
    x = NULL,
    y = "Number of occurences"
  ) +
  coord_flip()
```



# Sentiment Analysis Using AFINN Dictionary 

### Generate data frame with sentiment derived from the AFINN dictionary

```{r message=FALSE, warning=FALSE}
# creating df from AFINN dictionary
hp_afinn_sentiment_analysis <- get_sentiments("afinn") 
  

# joining by words from hp books that are also in the AFINN dictionairy

hp_afinn_sentiment_analysis <- inner_join(hp_words, hp_afinn_sentiment_analysis, by = "word") %>% 
  group_by(book, chapter) %>% 
  summarize(value = sum(value))
```

## Emotional arc of Harry Potter books


```{r echo=FALSE}

hp_afinn_sentiment_analysis %>% 
  ggplot(mapping = aes(x = chapter, y = value, fill = book)) +
  geom_col() +
  facet_wrap(facets = vars(book), scales = "free_x") +
  labs(
    title = "Emotional arc of Harry Potter books",
    subtitle = "AFINN sentiment dictionary",
    x = "Chapter",
    y = "Emotional score"
  ) +
  theme(legend.position = "none")
```


### Interpretation:

As most HP fans may expect, sentiment really shifts from mostly postitive to mostly negative as the series progresses. By Deathly Hollows, the sentiment score is negative almost the whole film. 

